\bf Hierarchical Network Structure Promotes Dynamical Robustness

### 2015-07-16 tex/figures.tex
update figure legend
move definition of hierarchy in terms of graph edit distance from SI to main text

\begin{figure*}[!htbp]
\centering
\noindent\includegraphics[width=1.0\textwidth]{fig/robustnessprocess.pdf}
\caption{{\bf Assessment of dynamical robustness.} (A) Any stochastic process induced on the stable and unstable regions of the space of fixed points can be represented as a state transition diagram. The conditional probability of remaining in the stable region in the context of such modifications, given a previously existing stable state, is then provided by estimating $R$. (B) The stochastic process that results from modifications to a given dynamical model over evolutionary timescales induces a stochastic process on the equivalence classes of fixed points $\mathcal{F}/{\sim}$ represented by the boxes in the top row. This in turn induces a process on the stable and unstable regions, which can be used to estimate robustness. We note that $R$ is distinct from the stationary probability of being within the stable region $p^{\infty}(S) = \frac{1-U}{2-R-U}$. (C) Each network architecture encoded by a directed graph, $\{G_1, \ldots, G_k \}$, that indicates the manner in which the variables of a biological network depend upon one another selects a particular subset of equivalence classes from $\mathcal{F}/{\sim}$ (see \ref{eq:jacgrapheqs}). Robustness can therefore be estimated independently for each network architecture.}
\label{fig:robustnessprocess}
\end{figure*}

@@

\begin{figure*}[!ht]
\centering
\noindent\includegraphics[width=1.0\textwidth]{fig/modsccsym.pdf}
\caption{{\bf Open systems, strongly connected components and symmetries of robustness.} (A) Example of the combination of open system modules to construct closed systems. (B) SCCs highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three variable systems. The most hierarchical network, top panel, is the one that maximizes the number of SCCs and the number of links between them. We therefore define hierarchy as $max(\hbox{ED}) - \hbox{ED}$ where ED is the edit distance representing the number of link addition/deletion operations necessary to transform a given graph into the most hierarchical one. The two panels in the middle represent examples of hierarchical modular systems that posess both modularity (i.e. SCCs with more than one variable) and hierarchy. (C) Symmetries of the $\hier$ transformation between graphs and SCCs. The transformation $a$ represents an interchange of SCCs, $b$ moving a link between nodes in a component and $c$ adding a link. All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only $a$ and $b$ are symmetries of robustness.}
\label{fig:modsccsym}
\end{figure*}

>>>>

\begin{figure*}[!htbp]
\centering
\noindent\includegraphics[width=1.0\textwidth]{fig/robustnessprocess.pdf}
\caption{{\bf Assessment of dynamical robustness.} (A) Any stochastic process induced on the stable and unstable regions of the space of fixed points can be represented as a state transition diagram. The conditional probability of remaining in the stable region in the context of such modifications, given a previously existing stable state, is then provided by estimating $R$. (B) The stochastic process that results from modifications to a given dynamical model over evolutionary timescales induces a stochastic process on the equivalence classes of fixed points $\mathcal{F}/{\sim}$ represented by the boxes in the top row. This in turn induces a process on the stable and unstable regions, which can be used to estimate robustness. We note that $R$ is distinct from the stationary probability of being within the stable region $P^{\infty}(S) = \frac{1-U}{2-R-U}$. (C) Each network architecture encoded by a directed graph, $\{G_1, \ldots, G_k \}$, that indicates the manner in which the variables of a biological network depend upon one another selects a particular subset of equivalence classes from $\mathcal{F}/{\sim}$ (see \ref{eq:jacgrapheqs}). Robustness can therefore be estimated independently for each network architecture.}
\label{fig:robustnessprocess}
\end{figure*}

@@

\begin{figure*}[!ht]
\centering
\noindent\includegraphics[width=1.0\textwidth]{fig/modsccsym.pdf}
\caption{{\bf Open systems, strongly connected components and symmetries of robustness.} (A) Example of the combination of open system modules to construct closed systems. (B) SCCs highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three variable systems. The most hierarchical network, top panel, is the one that maximizes the number of SCCs and the number of links between them.
% We therefore define hierarchy as $max(\hbox{ED}) - \hbox{ED}$ where ED is the edit distance representing the number of link addition/deletion operations necessary to transform a given graph into the most hierarchical one.
The two panels in the middle represent examples of hierarchical modular systems that posess both modularity (i.e. SCCs with more than one variable) and hierarchy. (C) Symmetries of the $\hier$ transformation between graphs and SCCs. The transformation $a$ represents an interchange of SCCs, $b$ moving a link between nodes in a component and $c$ adding a link. All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only $a$ and $b$ are symmetries of robustness.}
\label{fig:modsccsym}
\end{figure*}

### 2015-07-16 tex/methods.tex
move definition of hierarchy in terms of graph edit distance from SI to main text

\section{Hierarchy and Total ordering}\label{sec:totalordering}

@@

The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents. The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \reffigscc{} (top panel). This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components. The graph edit distance (ED) on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}. This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being maximally hierarchical. In this work we take $max(ED) - ED$ to be the definition of hierarchy, where $max(ED)$ is the maximum edit distance for all graphs with a given number of nodes.

>>>>

\section{Graph associated to the total ordering}\label{sec:totalordering}

@@

### 2015-07-16 tex/overview.tex
move definition of hierarchy in terms of graph edit distance from SI to main text

>>>>

\subsection{Quantification of network hierarchy}

The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents. The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \reffigscc{} (top panel). This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components. The graph edit distance (ED) on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}. This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being maximally hierarchical. In this work we take $max(ED) - ED$ to be the definition of hierarchy, where $max(ED)$ is the maximum edit distance for all graphs with a given number of nodes.
